---
title: "Task 4 - Permutation"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(readxl)
library(boot)
library(bootstrap)
library(ash)
library(fitdistrplus)
library(mixtools)

set.seed(12345)

salary_df <- read_csv("Salary_Data.csv")
```


Permutation


We can see the distribution of the Age of Experience here.

```{r histogram sal, echo=FALSE}
hist(salary_df$YearsExperience, breaks = "Sturges", freq = FALSE, main = "Histogram of Years of Experience", xlab="Years of Experience")
```


Seeing as there are two peaks, the distribution might be a mixture. Let's fit two normal distributions.

```{r mixtools, echo=FALSE}

exp <- normalmixEM(salary_df$YearsExperience, fast = TRUE)
```

```{r}
plot(exp, density = TRUE)
```

```{r mixture capture stats}
cat("Mixture Log Likelihood:", exp$loglik,"\n")
cat("First Normal Distribution proportion:", exp$lambda[1], ", Mean:",exp$mu[1], ", and Standard Deviation:",exp$sigma[1]  )
cat("\nSecond Normal Distribution proportion:", exp$lambda[2], ", Mean:",exp$mu[2], ", and Standard Deviation:",exp$sigma[2]  )
```

Now, let's fit try fitting just one theoretical distribution

```{r describe distribution}
descdist(salary_df$YearsExperience, discrete = FALSE)
```

```{r fitting one distribution}
fit.weibull <- fitdist(salary_df$YearsExperience, "weibull")
fit.norm <- fitdist(salary_df$YearsExperience, "norm")

```


```{r plot fitted distributions}
plot(fit.weibull)
plot(fit.norm)
hist(salary_df$Salary, breaks = "Sturges", freq = FALSE)
```

```{r aic of fit}
cat("Fitted Weibull AIC:",fit.weibull$aic )
cat("Fitted Normal AIC:", fit.norm$aic )
```



The Weibull distribution seems to explain our sample better, in the case of fitting one distribution.


The estimated parameters are:
```{r}
fit.weibull$estimate
```



## Weibull Permutation Testing

```{r creating weibull sample}
weibull.samp <- rweibull(30, shape = fit.weibull$estimate["shape"], scale = fit.weibull$estimate["scale"])

```


```{r weibull ks test, echo=FALSE}
weibull.ks.test.initial <- ks.test(salary_df$YearsExperience, weibull.samp)

cat("The p-value of the initial Kolmogorov-Smirnov test is:", weibull.ks.test.initial$p.value)
weibull.ks.test.D0 <- weibull.ks.test.initial$statistic
cat("\nThe Test Statistic is:", weibull.ks.test.D0)
```



```{r weibull ks permutation, include=FALSE}

B <- 10000
Z <- c(salary_df$YearsExperience, weibull.samp)
K <- 1:60
D <- numeric(B) 

options(warn = -1)

for (i in 1:B) {

k <- sample(K, size = 30, replace = FALSE)
x1 <- Z[k]
y1 <- Z[-k] #complement of x1
D[i] <- ks.test(x1, y1, exact = FALSE)$statistic
}

p <- mean(c(weibull.ks.test.D0, D) >= weibull.ks.test.D0)
options(warn = 0)
```
```{r, echo=FALSE}
cat("The obtained ASL is:", p)
hist(D, freq = FALSE, main = "Histogram of fitted Weilbull vs Original Sample KS test statistic", xlab="test statistic")
```


## Mixed Normal Distribution Permutation Testing

```{r creating mix norm sample}

N <- 30
sample_idx <- sample(1:2, prob = exp$lambda, size = N, replace = TRUE)
mixnorm.samp <- rnorm(N, mean = exp$mu[sample_idx], sd = exp$sigma[sample_idx])

```
```{r ks test mix norm}
mixnorm.ks.test.initial <- ks.test(salary_df$YearsExperience, mixnorm.samp)

cat("The p-value of the initial Kolmogorov-Smirnov test is:", mixnorm.ks.test.initial$p.value)
mixnorm.ks.test.D0 <- mixnorm.ks.test.initial$statistic
cat("\nThe Test Statistic is:", mixnorm.ks.test.D0)
```

```{r mix norm perm test, include=FALSE}
B <- 10000
Z <- c(salary_df$YearsExperience, mixnorm.samp)
K <- 1:60
D <- numeric(B) 

options(warn = -1)

for (i in 1:B) {

k <- sample(K, size = 30, replace = FALSE)
x1 <- Z[k]
y1 <- Z[-k] #complement of x1
D[i] <- ks.test(x1, y1, exact = FALSE)$statistic
}

p <- mean(c(mixnorm.ks.test.D0, D) >= mixnorm.ks.test.D0)
options(warn = 0)
```
```{r, echo=FALSE}
cat("The obtained ASL is:", p)
hist(D, freq = FALSE, main = "Histogram of fitted Normal Mixture vs Original Sample KS test statistic", xlab="test statistic")
```